---
title: "Macro effects"
author: "Ľuboš Perniš"
date: "09/04/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source('functions/helper_Error_calc.R')
source('functions/causalMatchFNN.R')
```

# Macro effects

We start again by creating the initial and the target location.

```{r data generation}

library(causalTree)

### First create data 
# This helper function generates y1
generate_y1 <- function(treatment_function) {
  return(y0 + eval(parse(text  = treatment_function)))
}

treatment_function <- '2 + 10 * x1'

## Create the synth dataset for D = 0 
# Create covariate x1 
set.seed(123) # For reproducibility
uniform <- runif(n = 400, min = 0, max = 10)
set.seed(123)
normal <- rnorm(n = 100, mean = 8)
x1 <- c(uniform, normal)
x1 <- x1[x1 <= 10 & x1 >= 0]
# Create y0 
y0 <- x1
# Define treatment
y1 <- generate_y1(treatment_function)

# Bind together to create simulation.1 d= 0 dataset
d0 <- as.data.frame(cbind(x1, y0, y1))

## Create the synth dataset for D = 1 
# Create covariate x1 
set.seed(1234) # For reproducibility
uniform <- runif(n = 400, min = 0, max = 10)
set.seed(1234)
normal <- rnorm(n = 100, mean = 2)
x1 <- c(uniform, normal)
x1 <- x1[x1 <= 10 & x1 >= 0]
# Create y0 
y0 <- x1
# Define treatment
y1 <- generate_y1(treatment_function)

# Bind together to create simulation.1 d= 0 dataset
d1 <- as.data.frame(cbind(x1, y0, y1))

# Clean the workspace
rm(x1, y1, y0, uniform, treatment_function, normal, generate_y1)

### Assign treatment to D = 0
set.seed(123)
random <- sample(1:nrow(d0))
treat_rows <- random[1:floor(0.5*length(random))]
d0$t <- NA
d0$t[treat_rows] <- 1
d0$t[-treat_rows] <- 0

### Create new var y (realised outcome)
d0$y <- ifelse(d0$t == 1, d0$y1, d0$y0)

# Clean the workspace
rm(random, treat_rows)
```

## Before macro

We show the true $\tau_1$
```{r}
mean(d1$y1-d1$y0)
```


We can make accurate predictions of $\tau_1^{PRED}$
```{r include=FALSE}
# Using matching
tauPred_m <- causalMatchFNN(d1, d0, 'x1')

# Using Forest
cf <-  causalForest(y ~ x1,
                     data=d0,
                     treatment=d0$t, 
                     split.Rule="CT",
                     split.Honest=T,
                     split.Bucket=F,
                     bucketNum = 5,
                     bucketMax = 100,
                     cv.option="CT",
                     cv.Honest=T,
                     minsize = 2L, 
                     split.alpha = 0.5,
                     cv.alpha = 0.5,
                     sample.size.total = floor(nrow(d0) / 2),
                     sample.size.train.frac = .5,
                     mtry = ceiling(ncol(d0)/3),
                     nodesize = 3,
                     num.trees= 5,
                     ncolx=length(grep('x', colnames(d0))),
                     ncov_sample= length(grep('x', colnames(d0)))
                     )

tauPredc <- mean(predict(cf, d1))
```
Using causal match `r tauPred_m` and using forest `r tauPredc`.

And here are the important corresponding errors
```{r}
# matching
calc_SE(d1, tauPred_m)
# forest
calc_SE(d1, tauPredc)
# naive error
calc_NPE(d0, d1)
```

## Changing the treatment effect function
Now we can change the treatment effect function to accommodate $Z$

$$
\theta_{i,j} = 2 + 10 * x^1 + 1* Z_j
$$

We start by creating a new variable $z^1$ with values 5 for initial location and 20 for the target location. We do this by replicating the entire data generation process.

```{r data generation with big difference in Z}
# This helper function generates y1
generate_y1 <- function(treatment_function) {
  return(y0 + eval(parse(text  = treatment_function)))
}

treatment_function <- '2 + 10 * x1 + z1'

## Create the synth dataset for D = 0 
# Create covariate x1 
set.seed(123) # For reproducibility
uniform <- runif(n = 400, min = 0, max = 10)
set.seed(123)
normal <- rnorm(n = 100, mean = 8)
x1 <- c(uniform, normal)
x1 <- x1[x1 <= 10 & x1 >= 0]
# Create y0 
y0 <- x1
# Create z1
z1 <- 5
# Define treatment
y1 <- generate_y1(treatment_function)

# Bind together to create simulation.1 d= 0 dataset
d0 <- as.data.frame(cbind(x1, z1, y0, y1))

## Create the synth dataset for D = 1 
# Create covariate x1 
set.seed(1234) # For reproducibility
uniform <- runif(n = 400, min = 0, max = 10)
set.seed(1234)
normal <- rnorm(n = 100, mean = 2)
x1 <- c(uniform, normal)
x1 <- x1[x1 <= 10 & x1 >= 0]
# Create y0 
y0 <- x1
# Create z1
z1 <- 20
# Define treatment
y1 <- generate_y1(treatment_function)

# Bind together to create simulation.1 d= 0 dataset
d1 <- as.data.frame(cbind(x1, z1, y0, y1))

# Clean the workspace
rm(x1, y1, y0, uniform, treatment_function, normal, generate_y1)

### Assign treatment to D = 0

set.seed(123)
random <- sample(1:nrow(d0))
treat_rows <- random[1:floor(0.5*length(random))]
d0$t <- NA
d0$t[treat_rows] <- 1
d0$t[-treat_rows] <- 0

### Create new var y (realised outcome)
d0$y <- ifelse(d0$t == 1, d0$y1, d0$y0)

# Clean the workspace
rm(random, treat_rows)

```

Now we can inspect the true ATE first $\tau_1$

```{r}
mean(d1$y1-d1$y0)
```
We also look at NPE which is very low in this case

```{r}
calc_NPE(d0, d1)
```

### Making predictions

```{r include=FALSE}
tauPred_m_z <- causalMatchFNN(d1, d0, 'x1')

# Using Forest
cf_z <-  causalForest(y ~ x1,
                     data=d0,
                     treatment=d0$t, 
                     split.Rule="CT",
                     split.Honest=T,
                     split.Bucket=F,
                     bucketNum = 5,
                     bucketMax = 100,
                     cv.option="CT",
                     cv.Honest=T,
                     minsize = 2L, 
                     split.alpha = 0.5,
                     cv.alpha = 0.5,
                     sample.size.total = floor(nrow(d0) / 2),
                     sample.size.train.frac = .5,
                     mtry = ceiling(ncol(d0)/3),
                     nodesize = 3,
                     num.trees= 5,
                     ncolx=length(grep('x', colnames(d0))),
                     ncov_sample= length(grep('x', colnames(d0)))
                     )

tauPredc_z <- mean(predict(cf, d1))
```

### Predictions
Our predictions $\tau_1^{PRED$:
- Causal Match: `r tauPred_m_z`
- Causal Forest: `r tauPredc_z`

```{r}
# Matching
calc_SE(d1, tauPred_m_z)
# Causal Forest
calc_SE(d1, tauPredc_z)
```

```{r Create first row of the table}
row1 <- data.frame(
  tau_1 = mean(d1$y1-d1$y0), 
  se_forest = calc_SE(d1, tauPredc_z), 
  se_match = calc_SE(d1, tauPred_m_z), 
  NPE = calc_NPE(d0, d1)
)


```



This is way worse than NPE

## Lowering the difference in $z^1$
We start as usual by creating our data. We assume the treatment function defined above. However, we change the value of $z^1$ to 5 for the intitial location and to 6 for the target location.


```{r}
# This helper function generates y1
generate_y1 <- function(treatment_function) {
  return(y0 + eval(parse(text  = treatment_function)))
}

treatment_function <- '2 + 10 * x1 + z1'

## Create the synth dataset for D = 0 
# Create covariate x1 
set.seed(123) # For reproducibility
uniform <- runif(n = 400, min = 0, max = 10)
set.seed(123)
normal <- rnorm(n = 100, mean = 8)
x1 <- c(uniform, normal)
x1 <- x1[x1 <= 10 & x1 >= 0]
# Create y0 
y0 <- x1
# Create z1
z1 <- 5
# Define treatment
y1 <- generate_y1(treatment_function)

# Bind together to create simulation.1 d= 0 dataset
d0 <- as.data.frame(cbind(x1, z1, y0, y1))

## Create the synth dataset for D = 1 
# Create covariate x1 
set.seed(1234) # For reproducibility
uniform <- runif(n = 400, min = 0, max = 10)
set.seed(1234)
normal <- rnorm(n = 100, mean = 2)
x1 <- c(uniform, normal)
x1 <- x1[x1 <= 10 & x1 >= 0]
# Create y0 
y0 <- x1
# Create z1
z1 <- 4
# Define treatment
y1 <- generate_y1(treatment_function)

# Bind together to create simulation.1 d= 0 dataset
d1 <- as.data.frame(cbind(x1, z1, y0, y1))

# Clean the workspace
rm(x1, y1, y0, uniform, treatment_function, normal, generate_y1)

### Assign treatment to D = 0

set.seed(123)
random <- sample(1:nrow(d0))
treat_rows <- random[1:floor(0.5*length(random))]
d0$t <- NA
d0$t[treat_rows] <- 1
d0$t[-treat_rows] <- 0

### Create new var y (realised outcome)
d0$y <- ifelse(d0$t == 1, d0$y1, d0$y0)

# Clean the workspace
rm(random, treat_rows, z1)

```

Now we can inspect the true ATE first $\tau_1$

```{r}
mean(d1$y1-d1$y0)
```
We also look at NPE which is very low in this case

```{r}
calc_NPE(d0, d1)
```

### Making predictions

```{r include=FALSE}
tauPred_m_z <- causalMatchFNN(d1, d0, 'x1')

# Using Forest
cf_z <-  causalForest(y ~ x1,
                     data=d0,
                     treatment=d0$t, 
                     split.Rule="CT",
                     split.Honest=T,
                     split.Bucket=F,
                     bucketNum = 5,
                     bucketMax = 100,
                     cv.option="CT",
                     cv.Honest=T,
                     minsize = 2L, 
                     split.alpha = 0.5,
                     cv.alpha = 0.5,
                     sample.size.total = floor(nrow(d0) / 2),
                     sample.size.train.frac = .5,
                     mtry = ceiling(ncol(d0)/3),
                     nodesize = 3,
                     num.trees= 5,
                     ncolx=length(grep('x', colnames(d0))),
                     ncov_sample= length(grep('x', colnames(d0)))
                     )

tauPredc_z <- mean(predict(cf_z, d1))
```

### Predictions
Our predictions $\tau_1^{PRED$:
- Causal Match: `r tauPred_m_z`
- Causal Forest: `r tauPredc_z`

```{r}
# Matching
calc_SE(d1, tauPred_m_z)
# Causal Forest
calc_SE(d1, tauPredc_z)
```

```{r}
row2 <- data.frame(
  tau_1 = mean(d1$y1-d1$y0), 
  se_forest = calc_SE(d1, tauPredc_z), 
  se_match = calc_SE(d1, tauPred_m_z), 
  NPE = calc_NPE(d0, d1)
)
```



## Explanations

1. Variance explained by $z^1$

```{r}
bindloc <- rbind(d0[, c('x1', 'z1', 'y0', 'y1')], d1[, c('x1', 'z1', 'y0', 'y1')])
theta <- 2 + 10 * bindloc$x1 + bindloc$z1
summary(lm(data=bindloc, theta ~ x1))


library(lme4)
multi <- lme4::lmer(data = bindloc, theta ~ x1 + (1|z1))
summary(multi)
ranef(multi)
```

